### 2019-07-29 reports/thesis/chapters/3_compoundPoisson.tex
Compound Poisson chapter

A special case of the compound Poisson is when $U\sim\gammaDist\left(\alpha,\beta\right)$ where $\alpha>0$ is the gamma shape parameter and $\beta>0$ is the gamma rate parameter. This was used for example in \cite{xu2009electronic}. This is known as the compound Poisson-gamma distribution. This is denoted by $X\sim\CPoisson(\lambda,\alpha,\beta)$ and has p.d.f.
\begin{equation}
  p_X(x) = 
  \begin{cases}
    \delta(x) \euler^{-\lambda} & \text{ for } x=0 \\ 
    \sum_{y=1}^{\infty}\frac{\beta^{y\alpha}}{\Gamma(y\alpha)}x^{y\alpha-1}\euler^{-\beta x}\euler^{-\lambda}\frac{\lambda^y}{y!} & \text{ for } x>0
  \end{cases}
  \ .
  \label{eq:compoundPoisson_pdf}
\end{equation}
Recall that $X|Y=\sum_{i=1}^YU_i$ which involves a sum of gamma random variables. It can be shown that
\begin{equation}
  X|Y\sim\gammaDist\left(Y\alpha,\beta\right).
\end{equation}

@@

It can be shown that the compound Poisson-gamma distribution is in the exponential family for fixed $\alpha$ \citep{jorgensen1987exponential}. To show this, it is useful to parametrise the compound Poisson-gamma distribution using the following:
\begin{equation}
  p=\frac{2+\alpha}{1+\alpha}
  \ ,
\end{equation}
\begin{equation}
  \mu=\frac{\lambda\alpha}{\beta}
  \ ,
\end{equation}
\begin{equation}
  \phi = \frac{\alpha+1}{\beta^{2-p}(\lambda\alpha)^{p-1}}
  \ .
\end{equation}
The parameters $p$, $\mu$ and $\phi$ are called the index, mean and dispersion parameters respectively. The parameters take the values of $1<p<2$, $\mu>0$ and $\phi>0$. After parametrising, it can be shown that the p.m.f.~at zero is
\begin{equation}
  \prob(X=0) = \exp
  \left[
      -\frac{\mu^{2-p}}{\phi(2-p)}
  \right]
\end{equation}
and the p.d.f.~for $x>0$ is
\begin{equation}
  p_X(x) = 
  \exp\left[
    \frac{1}{\phi}
    \left(
      x\frac{\mu^{1-p}}{1-p}-\frac{\mu^{2-p}}{2-p}
    \right)
  \right]
  \frac{1}{x}
  \sum_{y=1}^{\infty}W_y(x,p,\phi)
\end{equation}
where
\begin{equation}
  W_y = W_y(x,p,\phi)=\frac{x^{y\alpha}}{\phi^{y(1+\alpha)}(p-1)^{y\alpha}(2-p)^yy!\Gamma(y\alpha)}
  \ .
\end{equation}
The derivation is shown in Appendix \ref{chapter:appendix_tweedie}. This is in the form of a generalised linear model \citep{nelder1972generalized, nelder1972generalized_2, mccullagh1984generalized} for fixed $p$ or $\alpha$ because the above is in the form of a distribution in the dispersive exponential family. This has applications in, for example, insurance claim data \citep{jorgensen1994fitting, smyth2002fitting}.

@@

One special property of the compound Poisson-gamma distribution is that it is in the Tweedie dispersion exponential family \citep{jorgensen1987exponential}. It can be shown that it has a special variance mean relationship
\begin{equation}
  \variance[X] = \phi \mu^p
\end{equation}
where $1<p<2$. This is derived in Appendix \ref{chapter:appendix_tweedie}. It should be noted that this relationship is for fixed $p$. This is different from the linear variance mean relationship found at the start of the chapter which was for fixed $\alpha$ from assuming no beam hardening.

>>>>

A special case of the compound Poisson is when $U\sim\gammaDist\left(\alpha,\beta\right)$ where $\alpha>0$ is the gamma shape parameter and $\beta>0$ is the gamma rate parameter. This was used for example in \cite{xu2009electronic}. This is known as the compound Poisson-gamma distribution. This is denoted by $X\sim\CPoisson(\lambda,\alpha,\beta)$ and has p.d.f.
\begin{equation}
  p_X(x) = 
  \begin{cases}
    \delta(x) \euler^{-\lambda} & \text{ for } x=0 \\ 
    \displaystyle\sum_{y=1}^{\infty}\dfrac{\beta^{y\alpha}}{\Gamma(y\alpha)}x^{y\alpha-1}\euler^{-\beta x}\euler^{-\lambda}\frac{\lambda^y}{y!} & \text{ for } x>0
  \end{cases}
  \ .
  \label{eq:compoundPoisson_pdf}
\end{equation}
Recall that $X|Y=\sum_{i=1}^YU_i$ which involves a sum of gamma random variables. It can be shown that
\begin{equation}
  X|Y\sim\gammaDist\left(Y\alpha,\beta\right).
\end{equation}

@@

It can be shown that the compound Poisson-gamma distribution is in the exponential family for fixed $\alpha$ \citep{jorgensen1987exponential}. To show this, it is useful to parametrise the compound Poisson-gamma distribution using the following:
\begin{equation}
  p=\frac{2+\alpha}{1+\alpha}
  \ ,
\end{equation}
\begin{equation}
  \mu=\frac{\lambda\alpha}{\beta}
  \ ,
\end{equation}
\begin{equation}
  \phi = \frac{\alpha+1}{\beta^{2-p}(\lambda\alpha)^{p-1}}
  \ .
\end{equation}
The parameters $p$, $\mu$ and $\phi$ are called the index, mean and dispersion parameters respectively. The parameters take the values of $1<p<2$, $\mu>0$ and $\phi>0$. After parametrising, it can be shown that the p.m.f.~at zero is
\begin{equation}
  \prob(X=0) = \exp
  \left[
      -\frac{\mu^{2-p}}{\phi(2-p)}
  \right]
\end{equation}
and the p.d.f.~for $x>0$ is
\begin{equation}
  p_X(x) = 
  \exp\left[
    \frac{1}{\phi}
    \left(
      x\frac{\mu^{1-p}}{1-p}-\frac{\mu^{2-p}}{2-p}
    \right)
  \right]
  \frac{1}{x}
  \sum_{y=1}^{\infty}W_y(x,p,\phi)
\end{equation}
where
\begin{equation}
  W_y = W_y(x,p,\phi)=\frac{x^{y\alpha}}{\phi^{y(1+\alpha)}(p-1)^{y\alpha}(2-p)^yy!\Gamma(y\alpha)}
  \ .
\end{equation}
The derivation is shown in Appendix \ref{chapter:appendix_tweedie}. This is in the form of a generalised linear model \citep{nelder1972generalized, nelder1972generalized_2, mccullagh1984generalized} for fixed $p$ and $\phi$ because the above is in the form of a distribution in the dispersive exponential family. This has applications in, for example, insurance claim data \citep{jorgensen1994fitting, smyth2002fitting}.

@@

One special property of the compound Poisson-gamma distribution is that it is in the Tweedie dispersion exponential family \citep{jorgensen1987exponential}. It can be shown that it has a special variance mean relationship
\begin{equation}
  \variance[X] = \phi \mu^p
\end{equation}
where $1<p<2$. This is derived in Appendix \ref{chapter:appendix_tweedie}. It should be noted that this relationship is for fixed $p$ and $\phi$. This is different from the linear variance mean relationship found at the start of the chapter which was for fixed $\alpha$ and $\beta$ from assuming no beam hardening.
